Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-03-29 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@486 3fc257bb-8550-0410-81b3-a5c103da83f7

Discussion of PageRank and its matvec kernel.

@@

Discussion of triangle finding algorithm.

@@

>>>>

The PageRank algorithm assings a relative numeric rank to each
vertex in a graph.

It models the web as a directed graph $G(V,E)$, with each vertex $v
\in V$ representing a web page and each edge $e_{ij} \in E$
representing a hyperlink from $v_i$ to $v_j$.  The probability of
moving from $v_i$ to another vertex $v_j$ is $\alpha/d_{out}(v_i) +
(1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
0.8-0.9), $d_{out}(v)$ is the outdegree of vertex $v$, and $|V|$ is
the cardinality of $V$.  The first term represents the probability of
following a given link on page $v_i$; the second represents the
probability of moving to a random page.  For pages with no outlinks,
the first term is $\alpha/|V|$, indicating equal likelihood to move to
any other page.  Equivalently, the graph can be represented by a
matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
\alpha/d_{out}(v_i)$ if vertex $v_i$ links to $v_j$.  The PageRank
algorithm, then, is simply a power-method iteration in which the
dominating computation is matrix-vector multiplication $A^T x=y$,
where $x$ is the PageRank vector from the previous iteration.

The algorithm in Figure \ref{fig:pr} performs these
iterations.

@@

 \label{fig:pr}
\end{figure}

The MapReduce implementation performs two {\it map()} operations to
initialize the graph matrix $A$ and PageRank vector $x$.  A {\it
collate()} operation gathers all row entries $a_{ij}$ with their
associated $x_i$ entry, and a {\it reduce()} computes $a_{ij} x_i$.  A
second {\it collate()} gathers, for each $j$, all contributions to the
column sum $\sum a_{ij} x_i$, which is computed by a second {\it
reduce()}.  MPI\_Allreduce calls are used to compute global norms and
residuals.

A triangle in a graph is a clique of size three.  The algorithm in
Figure \ref{fig:tri} enumerates all such triplets of vertices (Vi, Vj,
Vk) where the three edges (Vi,Vj), (Vj,Vk), (Vi,Vk) exist in the
graph.  The input to the first map() is an undirected graph, a lower
triangular matrix from R-MAT generation, so that $V_i$ < $V_j$ for
every edge.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxxxxxx\=xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

Copy(): store $G_0$ = copy of KV pairs for input graph \\
Map(): \> {\bf Vertices} Convert edges to vertices \\
           \> \> input: Key = $(V_i,V_j)$, Value = NULL \\
           \> \> output: Key = $V_i$, Value = $V_j$ \\
           \> \> output: Key = $V_j$, Value = $V_i$ \\
Collate() \\
Reduce(): \> {\bf First Degree} add degree to one vertex in edge \\
              \> \> input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> for each value in MultiValue: \\
              \> \> output: Key = $(V_i,V_j)$, Value = $(D_i,0)$ if $Vi < Vj$ \\
              \> \> output: Key = $(V_j,V_i)$, Value = $(0,D_i)$ if $Vj < Vi$ \\
Collate() \\
Reduce(): \> {\bf Second Degree} add degree to second vertex in edge \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
              \> \> output: Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with Vi < Vj \\
Map(): \> {\bf Low Degree} vertex emits edge \\
           \> \> if $Di < Dj$, output: Key = $V_i$, Value = $V_j$ \\
           \> \> else if $Dj < Di$, output: Key = $V_j$, Value = $V_i$ \\
           \> \> else output: Key = $V_j$, Value = $V_i$ \\
Collate() \\
Reduce(): \> {\bf Angles} find angles for each central vertex $Vi$ \\
              \> \> input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> Loop $J$ over values in MultiValue: \\
	        \> \> Loop $K > J$ over values in MultiValue: \\
		  \> \> if $Vj < Vk$, output: Key = $(V_j,V_k)$, Value = $Vi$ \\
		  \> \> else output: Key = $(V_k,V_j)$, Value = Vi \\
Merge() \\ {\bf Add} KV pairs from original graph $G_0$ to Angles reduce() KV pairs \\
Collate() \\
Reduce(): \> {\bf Emit Triangles} \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = $(Vk, Vl, Vm, NULL, ...)$ \\
              \> \> if NULL exists in MultiValue, \\
	        \> \> output: Key = $(V_i,V_j,V_k)$, Value = NULL for each non-NULL value in MultiValue

@@

 \label{fig:tri}
\end{figure}

Further interations could find rectangles from open triangles, etc.
